---
title: "analysis"
format: html
---

## Motivation

From Chris Carlson in an email dated 05/23/25

> Is there a time where we could look at student La Vida and Discovery participation across 10-15 years, with some common academic/institutional metrics, but also variables that might be interesting to evaluate.

Completions:

-   Students who did La Vida before arrival, and those who did it in May of their first year after end of term

-   Discovery in Q1, Q2, or Q4

-   If they had to Re-Take Discovery?

**Main:**  

-   Overall FY Gordon GPA, and compared to their HS GPA 

-   FY retention rate (if they left, when did they participate?)

-   Persistence (if tjhey left before degree completion or not)

-   4Y and 6Y grad rate, never finished

**Comparisons:**

-   Financial Aid Tier

-   Family Resources or (previous) Estimated Family Contribution

-   Race/Ethnicity (white/non-white, if there’s not enough participants for analysis in segments)

-   Is/was an NCAA rostered athlete and by sport

-   By academic program or use current School groupings

-   Is there an Advancement giving or Alumni engagement metric?

-   Any way to tell if they married a Gordon alum? (Ha!)

-   And, what clever ones can you think of?

## Packages

```{r packages}
library(tidyverse)
```

## Data import

```{r data-import}

```

## Analysis

### Completions and course outcomes

#### Overall participation since 2010 (15 academic years)

```{r overall-participation}
lavida_disco_attempts
```
