---
title: "analysis"
format: html
---

## Motivation

From Chris Carlson in an email dated 05/23/25

> Is there a time where we could look at student La Vida and Discovery participation across 10-15 years, with some common academic/institutional metrics, but also variables that might be interesting to evaluate.

Completions:

-   Students who did La Vida before arrival, and those who did it in May of their first year after end of term

-   Discovery in Q1, Q2, or Q4

-   If they had to Re-Take Discovery?

**Main:**  

-   Overall FY Gordon GPA, and compared to their HS GPA 

-   FY retention rate (if they left, when did they participate?)

-   Persistence (if tjhey left before degree completion or not)

-   4Y and 6Y grad rate, never finished

**Comparisons:**

-   Financial Aid Tier

-   Family Resources or (previous) Estimated Family Contribution

-   Race/Ethnicity (white/non-white, if there’s not enough participants for analysis in segments)

-   Is/was an NCAA rostered athlete and by sport

-   By academic program or use current School groupings

-   Is there an Advancement giving or Alumni engagement metric?

-   Any way to tell if they married a Gordon alum? (Ha!)

-   And, what clever ones can you think of?

## Packages

```{r packages}
library(tidyverse)
```

## Data import

```{r data-import}
lavida_disco_attempts <- read_csv("../data/analysis_data/lavida_disco_attempts.csv")

course_evals <- read_csv("../data/analysis_data/course_evals.csv")
```

## Analysis

### Attempts, completions and course outcomes

#### Overall participation since 2010 (15 academic years)

```{r overall-participation}
lavida_disco_attempts |>
  filter(TRANSACTION_STS != "C") |> # current enrollments
  pull(ID_NUM) |>
  unique() |>
  length()
```

There have been 7,445 attempts at completing PED 015 and PED 016 made by 6,691 unique students.

#### Participation breakdown by PED 015 and PED 016

```{r relative-participation}
lavida_disco_attempts |>
  filter(YR_CDE < "2025") |>
  filter(ENTRY_YR >= 2010) |>
  mutate(ADV_REQ_CDE = if_else(ADV_REQ_CDE == "PED015", "Discovery",
    "Wilderness Expedition"
  )) |>
  mutate(ADV_REQ_CDE = factor(ADV_REQ_CDE,
    levels = c("Wilderness Expedition", "Discovery")
  )) |>
  ggplot(aes(y = 1)) +
  geom_bar(aes(fill = ADV_REQ_CDE),
    position = "fill"
  ) +
  geom_text(
    data = lavida_disco_attempts |>
      filter(ENTRY_YR >= 2010) |>
      group_by(ENTRY_YR, ADV_REQ_CDE) |>
      summarize(count = n()) |>
      mutate(
        prop = count / sum(count),
        perc = round(prop * 100, 0),
        PERC = str_c(perc, "%")
      ) |>
      filter(ADV_REQ_CDE == "PED016") |>
      select(ENTRY_YR, PERC),
    mapping = aes(x = 0, y = -0.5, label = PERC),
    color = "#006D22",
    fontface = "bold"
  ) +
  scale_fill_manual(values = c("#006D22", "#4D9964")) +
  facet_wrap(~ENTRY_YR, ncol = 5) +
  coord_polar(direction = -1) +
  ylim(c(-0.5, 2)) +
  labs(
    x = NULL, y = NULL, fill = NULL,
    title = "Proportionally more students choose Discovery in recent years",
    subtitle = str_wrap("Each facet corresponds to a particular incoming cohort. The percentage of students participating in the La Vida Wilderness Expedition (reported for each student cohort year) has been dropping since 2010.", 100)
  ) +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "top",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        axis.text = element_blank(),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/relative_participation_donuts.png",
       dpi = 600,
       height = 5.5,
       width = 6)
```

#### First-year Wilderness Expedition participation, before or after?

```{r fy-before-after}
lavida_disco_attempts |>
  filter(YR_CDE < "2025") |>
  filter(YR_LA_VIDA_TAKEN == "First") |>
  filter(ENTRY_YR >= 2010) |>
  mutate(PRE_POST_FY = factor(PRE_POST_FY,
    levels = c("Before", "After")
  )) |>
  ggplot(aes(y = 1)) +
  geom_bar(aes(fill = PRE_POST_FY),
    position = "fill"
  ) +
  geom_text(
    data = lavida_disco_attempts |>
      filter(YR_CDE < "2025") |>
      filter(YR_LA_VIDA_TAKEN == "First") |>
      filter(ENTRY_YR >= 2010) |>
      group_by(ENTRY_YR, PRE_POST_FY) |>
      summarize(count = n()) |>
      mutate(
        prop = count / sum(count),
        perc = round(prop * 100, 0),
        PERC = str_c(perc, "%")
      ) |>
      filter(PRE_POST_FY == "Before") |>
      select(ENTRY_YR, PERC),
    mapping = aes(x = 0, y = -0.5, label = PERC),
    color = gc_pal$gordon_blue,
    fontface = "bold"
  ) +
  scale_fill_manual(values = c(gc_pal$gordon_blue, gc_pal$scottie_cyan)) +
  facet_wrap(~ENTRY_YR, ncol = 5) +
  coord_polar(direction = -1) +
  ylim(c(-0.5, 2)) +
  labs(
    x = NULL, y = NULL, fill = NULL,
    title = str_wrap("Increasingly more students who participate in the Wilderness Expedition do so during the summer before their first year", 70),
    subtitle = str_wrap("Each facet corresponds to a particular incoming cohort. The percentage of students participating in the La Vida Wilderness Expedition before their first year (reported for each student cohort year) has been increasing since 2010. The reversal in the 2020 cohort is due to the COVID-19 pandemic.", 95)
  ) +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "top",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        axis.text = element_blank(),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/before_after_donuts.png",
       dpi = 600,
       height = 5.89,
       width = 6)
```

#### Discovery quad breakdown

```{r discovery-quad-breakdown}
lavida_disco_attempts |>
  filter(ADV_REQ_CDE == "PED015") |>
  filter(YR_CDE < "2025") |>
  drop_na(SUBTERM_CDE) |>
  ggplot(aes(x = SUBTERM_CDE)) +
  geom_bar(fill = "#4D9964") +
  facet_wrap(~YR_CDE, ncol = 5) +
  labs(x = NULL, y = "Attempts",
       title = "Discovery is most often attempted during the fall semester",
       subtitle = str_wrap("Each facet corresponds to a particular academic year. Given weather conditions, Discovery is not offered during the third quad, except in AY 2020 when a remote section was offered.", 95)) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90",
                                        linewidth = 0.25),
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "gray90",
                                  linewidth = 0.25),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/disco_quads.png",
       dpi = 600,
       height = 4.75,
       width = 6)
```

## Course outcomes

### Pass, fail, withdraw rates

This analysis is done directly in la_vida_report.qmd to render it as a sortable {gt} table.

### Course evaluations

La Vida and Discovery course evaluations from the previous three years with a comparison to other core classes commonly taken in the first two years.

#### Wilderness Expedition surveys

These surveys look at: (1) how has this class awakened your Christian faith?, (2) how would you rate this class overall?, (3) does La Vida continue to be a powerful experience in your life?

```{r la-vida-christian-faith-plot}
course_evals |>
  filter(group == "La Vida") |>
  count(awakening_christian_faith) |>
  ungroup() |>
  add_row(awakening_christian_faith = 1,
          n = 0) |>
  mutate(awakening_christian_faith = factor(awakening_christian_faith)) |>
  ggplot(aes(awakening_christian_faith, n)) +
  geom_col(fill = gc_pal$la_vida,
           width = 0.75) +
  annotate(geom = "segment",
           x = 1.415, xend = 4.45,
           y = -9.8, yend = -9.8,
           arrow = arrow(length = unit(0.15, "cm"),
                         type = "open"),
           linewidth = 0.3,
           lineend = "round",
           color = "gray30") +
  scale_x_discrete(labels = c("Not effective", "", "", "", "Extremely effective")) +
  coord_cartesian(ylim = c(0, 100), clip = "off") +
  labs(x = NULL, y = "Count",
       title = str_wrap("How effective is the course in awakening and strengthening your understanding of the Christian faith?", 75),
       subtitle = str_wrap("Based on 173 La Vida Wilderness Expedition course evaluations submitted over the last three academic years.", 95)) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90",
                                        linewidth = 0.25),
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "gray90",
                                  linewidth = 0.25),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/la_vida_faith.png",
       dpi = 600,
       height = 4,
       width = 6)
```

```{r la-vida-overall-plot}
course_evals |>
  filter(group == "La Vida") |>
  drop_na(overall_course) |>
  count(overall_course) |>
  ungroup() |>
  add_row(overall_course = 2,
          n = 0) |>
  mutate(overall_course = factor(overall_course)) |>
  ggplot(aes(overall_course, n)) +
  geom_col(fill = gc_pal$la_vida,
           width = 0.75) +
  scale_x_discrete(labels = c("Poor", "Weak", "Average", "Good", "Great")) +
  scale_y_continuous(breaks = seq(0, 125, 25)) +
  labs(x = NULL, y = "Count",
       title = "Overall, I rate this course as:",
       subtitle = str_wrap("Based on 173 La Vida Wilderness Expedition course evaluations submitted over the last three academic years.", 95)) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90",
                                        linewidth = 0.25),
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "gray90",
                                  linewidth = 0.25),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/la_vida_overall.png",
       dpi = 600,
       height = 4,
       width = 6)
```

```{r la-vida-powerful-experience}
course_evals |>
  filter(group == "La Vida") |>
  drop_na(la_vida_powerful_exp) |>
  count(la_vida_powerful_exp) |>
  mutate(la_vida_powerful_exp = if_else(la_vida_powerful_exp == 1,
                                        "No",
                                        "Yes")) |>
  ggplot(aes(la_vida_powerful_exp, n)) +
  geom_col(fill = gc_pal$la_vida,
           width = 0.75) +
  # scale_y_continuous(breaks = seq(0, 125, 25)) +
  labs(x = NULL, y = "Count",
       title = str_wrap("At this point, would you say La Vida [Wilderness Expedition] continues to be a powerful experience in your life?", 65),
       subtitle = str_wrap("Based on 173 La Vida Wilderness Expedition course evaluations submitted over the last three academic years.", 95)) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90",
                                        linewidth = 0.25),
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "gray90",
                                  linewidth = 0.25),
        strip.background = element_rect(fill = "gray90"),
        legend.position = "bottom",
        legend.key.size = unit(0.8, "lines"),
        plot.title.position = "plot",
        plot.tag = element_path(size = 2.8,
                                vjust = 0, hjust = 0,
                                alpha = 0.5),
        plot.tag.location = "plot",
        plot.tag.position = c(0, 0),
        text = element_text(family = "jost",
                            color = "gray30"),
        strip.text = element_text(face = "bold",
                                  color = "gray30"),
        plot.title = element_text(family = "neuton",
                                  size = 16),
        plot.subtitle = element_text(face = "italic",
                                     size = 10,
                                     color = "gray60"),
        plot.caption = element_text(face = "italic",
                                    size = 7,
                                    color = "gray60"))

ggsave("../plots/la_vida_powerful_exp.png",
       dpi = 600,
       height = 4,
       width = 6)
```

```{r course-eval-response-distributions}
course_evals |>
  ggplot(aes(x = awakening_christian_faith)) +
  geom_bar() +
  facet_wrap(~group, nrow = 1)
```

```{r course-evals-summary-comp}
course_evals |>
  group_by(group) |>
  summarize(across(.cols = awakening_christian_faith:overall_course,
                   .fns = list(mean_score = \(x) mean(x, na.rm = T),
                               se_score = \(x) sd(x, na.rm = T)))) |>
  view()
```
